```{r}
#| warning: false
#| message: false
library(tidyverse)
library(ggplot2)

raw_model_prices <- read_csv("data/price/WVS_prompts_v4_English - Model prices.csv")

raw_actual_tokens <- read_csv('data/price/openrouter-activity-tokens-20260519-164154.csv') %>% mutate(model = Slug) %>% select(-c(Slug))

raw_actual_price <- read_csv('data/price/openrouter-activity-spend-20260519-164107.csv') %>% mutate(model = Slug) %>% select(-c(Slug))

raw_requests = read_csv('data/price/openrouter-activity-requests-20260519-191304.csv') %>% mutate(model = Slug) %>% select(-c(Slug))

data <- read_csv("data/full_dataset.csv") %>% filter(answer != "")

prompts <- read_csv("prompts/processed_prompts.csv") %>% mutate(id = Original_ID, name = Name, ValidLengths = as.character(ValidLengths)) %>% select(id, name, ValidIntegers, ValidLengths)

data <- merge(data, prompts)

```

# Tokens

:::: columns
::: column

## Actual OpenRouter total tokens
```{r}

actual_tokens <- raw_actual_tokens %>% group_by(model) %>% summarize(total_input_tokens = sum(`Prompt Tokens`), total_output_tokens = sum(`Completion Tokens`+`Reasoning Tokens`), total_tokens = total_input_tokens + total_output_tokens)

actual_tokens %>% arrange(model) %>% mutate(across(-1, ~paste0(round(.x/1000000,2), "M")))
```

## OpenRouter average tokens per request

```{r}

requests <- raw_requests %>% group_by(model) %>% summarize(requests = sum(Requests))

tokens_requests <- merge(actual_tokens,requests)
tokens_requests

request_tokens <- tokens_requests %>% mutate(input_tokens = total_input_tokens/requests,
output_tokens = total_output_tokens/requests,total_tokens = total_tokens/requests) %>% select(model, input_tokens, output_tokens, total_tokens)

request_tokens %>% ungroup() %>% mutate(across(-1, ~paste0(round(.x/1000,2), "K")))

```

## OpenRouter average tokens per "survey"
```{r}
router_survey_tokens <- request_tokens %>% ungroup() %>% mutate(across(-1, ~.x*241))

router_survey_tokens %>% mutate(across(-1, ~paste0(round(.x/1000000,2), "M")))

```
## OpenRouter tokens for 100 surveys

```{r}
request_tokens  %>% ungroup() %>% mutate(across(-1, ~paste0(round(.x/1000000*241*100,2), "M")))
```

::: 
::: column

## Pydantic total tokens
```{r}

tokens <- data %>% filter(!is.na(input_tokens) & !is.na(output_tokens)) %>% group_by(id, model, temperature) %>% arrange(model,desc(temperature), id) %>% select(c(unique_id, model, id, temperature, input_tokens, output_tokens))

tokens <- tokens %>% group_by(model) %>% mutate(
total_input_tokens = sum(input_tokens),
total_output_tokens = sum(output_tokens),
total_tokens = total_input_tokens + total_output_tokens)
  
estimated_tokens <- tokens %>% select(model, total_input_tokens, total_output_tokens, total_tokens) %>% unique() %>% arrange(model)

estimated_tokens %>% ungroup() %>% mutate(across(-1, ~paste0(round(.x/1000000,2), "M")))

  
```

## Pydantic average tokens per question

```{r}

# Pydantic average tokens for each question

question_tokens<- tokens_prices %>% group_by(model, temperature, id) %>% mutate(
  avg_input_tokens = mean(input_tokens),
  avg_output_tokens = mean(output_tokens), avg_tokens = avg_input_tokens+avg_output_tokens) %>% select(-c(unique_id)) %>% arrange(model, id) %>% unique() %>% ungroup()

## Pydantic average of average tokens per question

avg_question_tokens <- question_tokens %>% group_by(model, temperature) %>% summarize(input_tokens = mean(avg_input_tokens), output_tokens = mean(avg_output_tokens), total_tokens = input_tokens + output_tokens)

avg_question_tokens %>% select(c(model, temperature, input_tokens, output_tokens, total_tokens)) %>% unique() %>% ungroup() %>% mutate(across(3:5, ~paste0(round(.x/1000,2), "K"))) 


```

## Pydantic tokens per survey

```{r}
survey_tokens <- avg_question_tokens %>% ungroup() %>% mutate(across(3:5, ~.x*241))
survey_tokens %>% ungroup() %>% mutate(across(-c(1,2), ~paste0(round(.x/1000000,2), "M")))
```

## Pydantic tokens per 100 surveys

```{r}

survey_tokens %>% ungroup() %>% mutate(across(-c(1,2), ~paste0(round(.x/1000000*100,2), "M")))

```

:::

::::


## Compare tokens
```{r}

renamed_actual_tokens <- actual_tokens %>% rename_with(~paste0('actual_', .), .cols = 2:ncol(.))
actual_tokens

renamed_estimated_tokens <- estimated_tokens %>% rename_with(~paste0('estimated_', .), .cols = 2:ncol(.))
estimated_tokens

compare_tokens <- merge(renamed_actual_tokens, renamed_estimated_tokens) %>% mutate(
  diff_input_tokens = 
  actual_total_input_tokens - 
      estimated_total_input_tokens, 
diff_output_tokens = 
  actual_total_output_tokens-
      estimated_total_output_tokens/1000000, 
diff_total_tokens = 
  diff_input_tokens + diff_output_tokens) %>% 
  select(c(model, starts_with('diff')))

compare_tokens %>% mutate(across(-1, ~paste0(round(.x/1000000,2), "M")))
```

# Prices

:::: columns
::: column

## OpenRouter total price

actual_total_price: the actual usage on OpenRouter

total_from_tokens: estimated usage actual OpenRouter reported token usage and the stated model prices.

Note: total from tokens is a little higher because I used the upper price bound for higher amount of tokens.

This shows that the providers are not lying about prices

```{r}

actual_price <- raw_actual_price %>% group_by(model) %>% summarize(actual_total_price = sum(Usage))

router_total_price <- merge(actual_price, actual_tokens) %>% merge(raw_model_prices) %>% mutate(across_models_price = sum(actual_total_price), price_from_input_tokens = total_input_tokens*input_price/1000000,
price_from_output_tokens= total_output_tokens*output_price/1000000,
total_from_tokens = price_from_input_tokens + price_from_output_tokens) %>% select(c(model, actual_total_price, total_from_tokens))

router_total_price %>% ungroup() %>% mutate(across(-1, ~paste0('$', round(.x,2))))

```

# Price per survey based on OpenRouter tokens

```{r}
router_survey_price <- merge(router_survey_tokens, raw_model_prices) %>% group_by(model) %>% mutate(survey_input_price = input_tokens*input_price/1000000, survey_output_price = output_tokens*output_price/1000000, survey_price = survey_input_price + survey_output_price) %>% select(c(model, survey_input_price, survey_output_price, survey_price))

router_survey_price %>% ungroup() %>% mutate(across(-1, ~paste0('$', round(.x,2))))

```

# Price for 100 surveys based on OpenRouter tokens

```{r}

router_survey_price %>% ungroup() %>% mutate(across(-1, ~paste0('$', round(.x*100,2))))

```

:::
::: column

## Total price based on Pydantic tokens

```{r}
tokens_prices <- merge(estimated_tokens, raw_model_prices) 
tokens_prices

total_price <- tokens_prices %>% group_by(model) %>% mutate(
    total_input_price = total_input_tokens*input_price/1000000,
 total_output_price = total_output_tokens*output_price/1000000,
total_price = total_input_price + total_output_price) %>% select(c(model, total_price)) %>% unique() %>% ungroup() %>% mutate(across_models_price = sum(total_price)) %>% arrange(model)

total_price %>% mutate(across(-1, ~paste0('$', round(.x,2))))

```

:::

## Price per survey based on Pydantic tokens

```{r}

pydantic_survey_price <- merge(survey_tokens, raw_model_prices) %>% group_by(model, temperature) %>%
  summarize(
    avg_input_price = sum(input_price*input_tokens/1000000), 
  avg_output_price = sum(output_price*output_tokens/1000000), 
  total_price = avg_input_price+avg_output_price)

pydantic_survey_price %>% ungroup() %>% mutate(across(3:5, ~paste0('$', round(.x,2))))

```

# Trials left

## Number of trials run per model per question

```{r}
# First, drop non-sense values (answer must contain integers in ValidIntegers with a length that is in ValidLengths)

validated <- data %>% mutate(answer = str_replace_all(answer, "[\\[\\]]", "")) %>% filter(!is.na(answer) & answer != "")

validated <- validated %>% mutate(answer_length = lengths(strsplit(answer, ","))) %>% filter(
    mapply(\(ans, vi) {
      all(as.integer(strsplit(ans, ",")[[1]]) %in% as.integer(strsplit(vi, ",")[[1]]))
    }, answer, ValidIntegers),
    mapply(\(len, vl) {
      as.character(len) %in% strsplit(vl, ",")[[1]]
    }, answer_length, ValidLengths)
  )


## Now count
count <- validated %>% group_by(model,name) %>% summarize(Count=n()) %>% group_by(model, Count) %>% summarize(Names = paste(name, collapse=", "))
count
```


## Minimum number of trials per model

```{r}
count %>% group_by(model) %>% summarize(MinCount = min(Count), MaxCount = max(Count), MinNeeded = 100-MinCount)
```

## Average minutes per 241 questions

```{r}

model_time <- data %>% filter(!is.na(time)) %>% 
  group_by(run_id, model) %>% 
  arrange(time) %>% 
  mutate(diff = lead(time) - time) %>% 
  select(c(model, run_id, diff))

#max_diff = max(model_time$diff, na.rm=TRUE)
#model_time <- model_time %>% filter(diff < max_diff) 

ggplot(model_time, aes(y=diff, x = model)) + geom_boxplot() + theme_classic()

```

```{r}

formatted_time <- function(min){
  if_else(min < 60, 
  paste(round(min,2), "min.", sep=" "), 
  paste(round(min/60,2), "hr.", sep = " "))
}

final_model_time <- model_time %>% ungroup() %>% group_by(model) %>% filter(!is.na(diff)) %>%
  filter(between(diff, quantile(diff, 0.05), quantile(diff, 0.95))) %>%
  summarize(avg_min = mean(diff, na.rm = TRUE)*241/60, avg_time = formatted_time(avg_min)) %>% mutate(time_10 = formatted_time(10*avg_min), time_100 = formatted_time(100*avg_min), total_time_100 = formatted_time(sum(avg_min*100))) %>% arrange(avg_min) %>% select(c(model, avg_time, time_10, time_100, total_time_100))

final_model_time

```